---
title: "replr: A Local JSON REPL"
author: "Package Maintainer"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{replr: A Local JSON REPL}
  %\VignetteEngine{rmarkdown::html_vignette}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

# Introduction

`replr` provides a lightweight HTTP server that executes R code. By default the
server returns the console text produced by the expression. Set `plain = FALSE`
to receive structured JSON instead. This vignette demonstrates the core
features for interacting with the server from R.

# Starting and Stopping the Server

Use `start_server()` to launch the HTTP server. By default it listens on port
8080. To run the server in the background, set `background = TRUE`.

```{r start-server, eval=FALSE}
# start the server on port 8123
replr::start_server(port = 8123, background = TRUE)
```

Check that the server is running with `server_status()`.

```{r status, eval=FALSE}
replr::server_status(port = 8123)
```

To stop the server, call `stop_server()`.

```{r stop-server, eval=FALSE}
replr::stop_server(port = 8123)
```

# Executing Code

`exec_code()` sends R expressions to the server for evaluation. The response is
a list that can include standard output, warnings, errors, plot paths and a
summary of the result.

```{r exec-code, eval=FALSE}
# returns console text
replr::exec_code("1 + 1", port = 8123)

# request JSON output
res <- replr::exec_code("1 + 1", port = 8123, plain = FALSE)
res$result_summary$type
## [1] "double"
```

## Controlling the Output

Several arguments let you tailor the response:

- `plain`: return console text (`TRUE`) or JSON (`FALSE`).
- `summary`: include a summary of the returned value.
- `output`, `warnings`, `error`: toggle these fields in the response.

To change how many rows of a data frame are shown in the summary preview, set
the global option `replr.preview_rows`:

```{r preview-option, eval=FALSE}
options(replr.preview_rows = 10)
```

Summaries can be disabled entirely:

```{r disable-summary, eval=FALSE}
replr::exec_code("mean(1:5)", port = 8123, summary = FALSE)
```

```{r custom-call, eval=FALSE}
# return only the calculated value as plain text (default)
replr::exec_code("sqrt(25)", port = 8123)

# JSON output
replr::exec_code("sqrt(25)", port = 8123, plain = FALSE)
```

# Converting to a Tibble

The response object has an S3 method `as_tibble()` which converts the result to
an easy-to-work-with tibble.

```{r as-tibble, eval=FALSE}
res <- replr::exec_code("summary(lm(mpg ~ cyl, data = mtcars))", port = 8123, plain = FALSE)

tibble <- as_tibble(res)
tibble
## # A tibble: 1 × 5
##   type  formula          r_squared aic  coefficients
##   <chr> <chr>                 <dbl> <dbl> <list>
## 1 lm    mpg ~ cyl             0.75    NA  <dbl [2,2]>
```

# Example Session

Below is a typical interactive session that showcases multiple commands.

```{r session-example, eval=FALSE}
library(replr)

start_server(port = 8123, background = TRUE)

# console text
exec_code("mean(1:5)", port = 8123)

# JSON result
exec_code("mean(1:5)", port = 8123, plain = FALSE)$result_summary$type
## [1] "double"
exec_code("plot(1:10)", port = 8123)
## $status
## [1] "success"
##
## $plots
## [1] "r_comm/images/plot_20250618_203333_4.png"
exec_code("warning('demo')", port = 8123)
## $warning
## [1] "demo"

stop_server(port = 8123)
```

# Conclusion

`replr` makes it easy to evaluate R code from external clients or scripts. By
default you get the printed console text, but JSON is available when needed.
This vignette covered starting the server, executing code with different
options, and converting results for further analysis.

